---
title: "Machine Learning Project - Cousera Week 4"
author: "Ridzuan Mohamad"
date: "5/11/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(kernlab)
library(randomForest)
library(rpart)
library(e1071)
```

## 1. Overview
The objectives of this project is to use the data from accelerometers on the belt, forearm, arm, and dumbbell obtain from 6 participants (“Pedro”, “Jeremy”, “Adelmo”, “Eurico”, “Carlitos” and “Charles”) and to “predict” the manner in which they did the exercise. The 6 participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The outcome of the exercises performed by the 6 participants, was recorded into 2 types of data sets, namely, "training" and "test" (“pml-training.csv” and “pml-testing.csv”) from the following URL link source:

1. Training Data: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
2. Test Data: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## 2. Data Exploratory Analysis

### 2.1 Data Loading

The training dan test dataset is provided by the data provider and we going to download both dataset from the URL link described above.

```{r}
# set the URL for the download
url_train_data <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_test_test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

file_training_set = file.path("dataset","pml-training.csv")
file_testing_set = file.path("dataset","pml-testing.csv")

# download the datasets
if (!file.exists(file_training_set) || !file.exists(file_testing_set)){
    if(!dir.exists(file.path("dataset"))){
        dir.create("dataset")
    }
    
    download.file(url_train_data, file_training_set, method="curl")
    download.file(url_test_test, file_testing_set, method="curl")
}

# load into data.frame
train.dataset.raw <- read.csv(file_training_set, na.strings = c("NA", ""))
test.dataset.raw <- read.csv(file_testing_set, na.strings = c("NA", ""))
```

### 2.2 Exploring and Cleansing

The downloaded dataset have the same variable which is `r dim(train.dataset.raw)[2]` and difference numbers of observation data. The training dataset have `r dim(train.dataset.raw)[1]` observations data and the test dataset `r dim(test.dataset.raw)[1]` have observations data. It is normal practices to have training dataset is larger than the test dataset.

There are (`r dim(train.dataset.raw)[2]`) variable (predictors) but we only need a few which will contribute to the accuracy of our prediction model later on. Refering to **_Appendix A - Dataset Structure for Raw Dataset_**, we found out there are a lot of variable contain only **NA**, **#DIV/0!** which is should be remove from the dataset. There are 7 variables is used for data identificatoin such as `X`, `user_name`, `raw_timestamp_part_1`, `raw_timestamp_part_2 `, `cvtd_timestamp `, `new_window` and `num_window` need to be remove as well.

```{r}
##Remove NA column & column 1-7 as identifier column
train.dataset.clean <- train.dataset.raw[, colSums(is.na(train.dataset.raw)) == 0]
train.dataset.clean <- train.dataset.clean[, -c(1:7)]

##Remove NA column & column 1-7 as identifier column
test.dataset.clean <- test.dataset.raw[, colSums(is.na(test.dataset.raw)) == 0]
test.dataset.clean <- test.dataset.clean[, -c(1:7)]
```

Now, we only have `r dim(train.dataset.clean)[2]` variable for both datasets and store it in the new data frame known as `train.dataset.clean` & `test.dataset.clean`. Detail dataset structure please refer to **_Appendix B - List of Variable for Clean Dataset_**

Exploring the classes data it shows that the _classe A_ have more records compare to other classes. Based on the **_Appendix C - Histrogram_** and the table below; _classe A_ have more than 1000 records compared to other classes.

```{r}
## display records into table by classe
table(train.dataset.clean$classe)
```

## 3. Building Prediction Model

The clean training dataset (_train.dataset.clean_) is split into two (2) partition :-

1. Training Set ( _trainset_ ) - 70%
2. Cross Validation Set ( _crossvalidationset_ ) - 30%

The Training set mention above will be fit into **Random Forest Algorithm** to create a prediction model which will be used to make a prediction. The prediction model information can be refered on **_Appendix D - Prediction Model Summary_** The other sub-training set ( _Cross Validation_ ) will be used to evaluate the prediction model accuracy.

### 3.1 Data Partioning

```{r}
inTrain <- createDataPartition(y=train.dataset.clean$classe, p=0.7, list = FALSE)
trainset <- train.dataset.clean[inTrain,]
crossvalidationset <- train.dataset.clean[-inTrain,]
```

### 3.2 Random Forest

```{r}
set.seed(1235)

## create proediction model using Random Forest Algorithm
model.fit <- randomForest(classe ~ ., data=trainset, method="class", importance=TRUE, 
                          proximity=TRUE, ntree=30)
rf.predict <- predict(model.fit, crossvalidationset)

## evaluate prediction model
confusion.matrix <- confusionMatrix(rf.predict, crossvalidationset$classe)
```

The cross validation check shows that the accuracy is `r format(confusion.matrix$overall[1]*100, digits=2)`%. The detail can be refered at **_Appendix E - Confusion Matrix Output_**.

## 4. Applying Model to Test Data

Fit the clean test dataset into the prediction model and print the output.

```{r}
final.predict <- predict(model.fit, test.dataset.clean)
final.predict
```

## References

Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers’
Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial
Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-
61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-
34459-6_6.

\newpage
---

#Appendixies

## Appendix A - Dataset Structure for Raw Dataset 

```{r}
str(train.dataset.raw)
```

## Appendix B - List of Variable for Clean Dataset 

```{r}
names(train.dataset.clean)
```

## Appendix C - Histogram 

```{r, warning=FALSE}
ggplot(data = train.dataset.clean, aes(x=classe)) +
    geom_histogram(stat = "count") +
    scale_x_discrete(name = "classe") +
    scale_y_continuous(name = "count") +
    ggtitle("Classes")
```

## Appendix D - Prediction Model Summary

```{r, warning=FALSE}
model.fit
```

## Appendix E - Confusion Matrix Output

```{r, warning=FALSE}
confusion.matrix
```